---
title: "Problem 3 (6 credits)"
subtitle: "HW2"
author: "Michael Brucek (mbrucek), Eduardo Chavez (echavezh), Kevin Grady (grady133), Danny Moncada (monca016), Tian Zhang (zhan7003)"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

```{r}
suppressWarnings(suppressPackageStartupMessages({
  library(TSA)
  library(forecast)
  library(ggplot2)
  library(dplyr)
}))
```

# Boston Crime Data Analysis

## Question 1

Please pull out the crime frequency data we got from Homework 1 - Problem 3 - Question 4. You may re-plot the time series to refresh yourself about the pattern.

```{r}
crime=read.table("crime.txt",header=T)

N=dim(crime)[1]
crime_aggr=aggregate(rep(1,N),list(year=crime[,1],month=crime[,2],day=crime[,3]),sum)
crime_aggr=crime_aggr[order(crime_aggr[,1],crime_aggr[,2],crime_aggr[,3]),]

Y_crime = crime_aggr[,4]
ts.plot(Y_crime, main = "Boston Crime (June 2015 - September 2018)", 
        xlab = "Time (daily)", ylab = "Crime Frequency")
```

### a) (1 credit)

First, let's fit an `auto.arima()` to find out a good ARIMA model for the data. Again, notice that, `auto.arima()` provides a "good" model but not necessarily the optimal. We will learn more concrete model selection techniques in Lecture 6.

**Hints**:

  - use `auto.arima()` function 

```{r}
auto_crime <- auto.arima(Y_crime)
auto_crime
```

### b) (2 credits)

What's the model? For example
$$
(Y_t - 10) = 0.4\cdot (Y_{t-1} - 10) + e_t - 0.8 \cdot e_{t-1}
$$

**Hints**:

  - The mean value comes with every $Y_t$. In the example above, the mean value is 10.
  
  - R assumes positive sign for MA models. In the example above, R would show -0.8, rather than +0.8 for the MA(1) coefficient

Please write down the model below:
$$
(Y_t - 270.6409) = 0.9888\cdot (Y_{t-1} - 270.6409) + e_t - 0.7142 \cdot e_{t-1} - 0.2542 \cdot e_{t-2} + 0.0446 \cdot e_{t-3}
$$

### c) (1 credit)

Are any of the coefficients significant?

**Hints**:

  - A coefficient is significant if its magnitude is (roughly) at least twice as large as its standard error.


```{r}
#Please write down your answer below
auto_crime
#All of the coefficients of the model except for ma3 (0.0446) are significant according 
#to the table above.  All of these coefficients have a magnitude twice as large as its
#standard error.

```

### d) (2 credits)

Please superimpose the fitted values on the original crime frequency time series. Does the model sufficiently explain the data?

**Hints:**

  - The fitted values can be calculated by `the original time series - arima_fit$residuals`

```{r}
crime_aggr$fitted <- Y_crime - auto_crime$residuals
t.test(Y_crime, crime_aggr$fitted)
#According to the two sample t-test results above, we cannot reject the null hypothesis 
#that the true difference in means is equal to 0. 

ts.plot(Y_crime, main = "Boston Crime with auto arima residuals fitted in red",
                    xlab = "Time (daily)", ylab = "Crime Frequency")
lines(crime_aggr$fitted, col = "red")

#Combining with the time series plot of the original data and the fitted series,
#we can safely say that the model generated via the auto arima process sufficiently 
#explains the Boston crime data.
```
