---
title: "Fixed Effects"
subtitle: "Simulating Fixed Effects"
author: "Danny Moncada monca016"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

```{r}
# Authors: Gordon Burtch and Gautam Ray
# Course: MSBA 6440
# Session: Fixed Effects
# Topic: Simulating Fixed Effects
# Lecture 5

suppressWarnings(suppressPackageStartupMessages({
library(stargazer)
library(plm)
library(car)
}))
```


```{r}

# Read in the data
CountryData<-read.csv("CountryData.csv")

# Create an ordinary least squares regression
ols<-lm(y ~ x1, data=CountryData)
summary(ols)


yhat <-ols$fitted
plot(CountryData$x1, CountryData$y, pch=19, xlab="x1", ylab="y")
abline(lm(CountryData$y~CountryData$x1),lwd=3, col="red")

fixed.dum <-lm(y ~ x1 + factor(country), data=CountryData)
summary(fixed.dum)

yhat<-fixed.dum$fitted
scatterplot(yhat~CountryData$x1|CountryData$country, boxplots=FALSE, xlab="x1", ylab="yhat",smooth=FALSE)
abline(lm(CountryData$y~CountryData$x1),lwd=3, col="red")

#Individual Effect Models

fixed <-plm(y ~ x1, data=CountryData, index=c("country", "year"), effect = "individual", model="within")
summary(fixed)

pool <-plm(y ~ x1, data=CountryData, index=c("country", "year"), model="pooling")
summary(pool)


pFtest(fixed, ols)
pFtest(fixed, pool)

random <-plm(y ~ x1, data=CountryData, index=c("country", "year"), effect = "individual", model="random")
summary(random)

phtest(fixed, random)

#Two-ways effect

fixed.twoways <-plm(y ~ x1, data=CountryData, index=c("country", "year"), effect ="twoways", model="within")
summary(fixed.twoways)

pFtest(fixed.twoways, ols)

pFtest(fixed.twoways, fixed)

fixed.time <-plm(y ~ x1 + factor(year), data=CountryData, index=c("country", "year"), model="within")
summary(fixed.time)

pFtest(fixed.time, fixed)

random.twoways <-plm(y ~ x1, data=CountryData, index=c("country", "year"), effect="twoways", model="random")
summary(random.twoways)

phtest(fixed.twoways, random.twoways)

# Lagrange Multiplier does a comparison with the Pooling model


plmtest(fixed, effect ="individual")
plmtest(fixed, effect ="time")
plmtest(fixed, effect ="twoways")