---
title: "DiD Regression"
author: "Danny Moncada monca016"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---


```{r}

suppressWarnings(suppressPackageStartupMessages({
library(dplyr)
library(ggplot2)
library(stargazer)
library(readr)
library(plm)
}))

```

```{r}


#### Load the data ####
MyData = read.csv("TSTV-Obs-Dataset.csv")


#how long is the period of observation?
max(MyData$week)-min(MyData$week)

#How many subjects got TSTV? (Treated)
length(unique(MyData$id[MyData$premium==TRUE]))

#How many subjects did not get TSTV? (Control)
length(unique(MyData$id[MyData$premium==FALSE]))

#In what 'week' does the "treatment" begin?
min(unique(MyData$week[MyData$after==TRUE]))



# As descriptive visualization, let's look at average weekly viewership for both premium and regular viewers
Week_Ave = MyData %>% group_by(week, premium) %>%
  summarise(ave_view = mean(view_time_total_hr)) %>% ungroup()
ggplot(Week_Ave, aes(x = week, y = ave_view, color = factor(premium))) + 
  geom_line() + 
  geom_vline(xintercept = 2227, linetype='dotted') + 
  ylim(0, 6) + xlim(2220,2233) + 
  theme_bw()



#### Difference in Differences Regression ####
# Interpret the treatment effect
did_basic = lm(log(view_time_total_hr+1) ~ premium*after, data=MyData)
summary(did_basic)


# Let's try replacing the treatment dummy with subject fixed effects.
# What happened to the estimate of premium?
did_fe = plm(log(view_time_total_hr+1) ~ premium*after, data = MyData, index=c("id"), effect="individual", model="within")
summary(did_fe)

# Further add week fixed effects
did_sfe_tfe = plm(log(view_time_total_hr+1) ~ premium*after, data = MyData, index=c("id", "week"), effect="twoway", model="within")
summary(did_sfe_tfe)


# Let's try dynamic DiD instead.
did_dyn_sfe_tfe <- lm(log(view_time_total_hr+1) ~ premium + factor(week) + premium*factor(week), data = MyData)
summary(did_dyn_sfe_tfe)

# Let's retrieve the coefficients and standard errors, and create confidence intervals
model = summary(did_dyn_sfe_tfe)
coefs_ses = as.data.frame(model$coefficients[16:28,c("Estimate", "Std. Error")])
colnames(coefs_ses) = c("beta", "se")
coefs_ses = coefs_ses %>%
  mutate(ub90 = beta + 1.96*se,
         lb90 = beta - 1.96*se,
         week = 1:nrow(coefs_ses))

# Let's connect the estimates with a line and include a ribbon for the CIs. 
ggplot(coefs_ses, aes(x = week, y = beta)) + 
  geom_line() + 
  geom_hline(yintercept=0,linetype="dashed") + 
  geom_vline(xintercept=6,linetype="dashed") + 
  geom_ribbon(aes(ymin = lb90, ymax = ub90), alpha = 0.3) + 
  theme_bw()


# Time for our placebo test... 
# Let's limit to pre-period data, and shift the treatment date back in time, artificially, and see if we see sig differences pre treatment.
# Again, recall first week when treatment starts
MyDataPre <- MyData[MyData$after==0,]
max(MyDataPre$week)
MyDataPre$after <- MyDataPre$week > 2224
did_log_basic_placebo <- lm(data=MyDataPre,log(view_time_total_hr+1)~premium+after+premium*after)
summary(did_log_basic_placebo)