---
title: "Problem 1 (6 percent)"
subtitle: "Workday 1"
author: "Danny Moncada and monca016"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

```{r}
suppressWarnings(suppressPackageStartupMessages({
  library(TSA)
  library(ggplot2)
  library(dplyr)
  library(forecast)
}))
```

# Simulations

## Question 1

Please simulate one sample path of a white noise process of length $T=200$ and save it into a time series object

* Please use name `e1` for the object that would contain the data

**Hints**:

1. **Always** specify a seed before generating random numbers. This way your results are reproducible!
1. Use `ts` to convert a vector to a Time-Series

```{r}

# your white noise sample path should be called "e1"
set.seed(42)

T <- 200L

y <- rnorm(T)
e1 <- ts(y)

```


## Question 2

Please plot the sample path $e_t$

**Hints**:

1. You can use `time(x)` to extract the time component of time series
2. Use `geom_line` if you are using `ggplot`

```{r}
etplot <- ggplot(data.frame(ts = time(e1), yt = e1), aes(x = ts, y = yt)) + geom_line(alpha = 0.5)
etplot
```


## Question 3

Please compute the mean and standard deviation of that process,
add them as horizontal lines to the plot (red for the mean and blue for 2 standard deviations around the mean)

**Hints**:

* `geom_hline` function may be helpful

```{r}

# please write your code below
mu = mean(e1)
sd = sd(e1)

etplot + geom_hline(yintercept = mu, color = "red") + geom_hline(yintercept = 2*sd, color = "blue") +
            geom_hline(yintercept = -2*sd, color = "blue")
```


## Question 4

Please generate a sample path from a random walk with the length of $T=200$ and plot it. 

* Please use the variable name $Y4$ for the array that would contain the sample path.

* Please **do not overwrite** variables created in the previous chunks (such as $e1$ that you created earlier)

**Hints**:

1. Try avoiding using loops. You don't really need a loop here.


```{r}
set.seed(42)

T <- 200L

# please write your code below
Y4 <- cumsum(rnorm(T))

y4plot <- ggplot(data.frame(ts = 1:T, yt = Y4), aes(x = ts, y = yt)) + geom_line(alpha = 0.5)
y4plot

```


## Question 5

Please generate 50 sample paths from a random walk with the length of $T=200$ and plot all of them using different colors.

* Please use the data.frame named `randwalk` that would contain all the sample paths in one long line called column Y. (That is, `randwalk$Y`)

**Hints**:

1. Use `dplyr` chains! Try avoiding using for-loops. 
2. At this point, you should be using ggplot
3. Create the data in the *long* format:
    - I would prefer if you first generate one very long sequence of white noise first, put into a data.frame and then add columns for sample path id and time 

```{r}
set.seed(42)

N <- 50L # 50 sample paths
T <- 200L 

# please write your code below
randwalk <- data.frame(Y = rep(rnorm(N*T)), 
                   id = rep(1:N, each = T), 
                   t = rep(1:T, N))

randwalk$Yt = do.call(c, tapply(randwalk$Y, randwalk$id, FUN = cumsum))

randwalk_plot <- ggplot(data = randwalk, aes(x = t, y = Yt, color = as.factor(id))) +
            geom_line() + theme(legend.position = "none")

randwalk_plot
```

## Question 6

Compute the mean function $E[Y_t]$ for every $t$ and add to your plot as well as the variance function $\mathrm{Var}[Y_t]$ for every $t$ and add (standard deviation) to your plot too

**Hints**:

1. Use `dplyr` chains! Try avoiding using for-loops. 
2. Create a separate data.frame with means and standard deviations that will help with extra layers in ggplot
3. At this point, you should be using ggplot

```{r}
randwalk2 <- randwalk %>%
                  group_by(t) %>%
                      mutate(mu = mean(Yt), 
                             upper = mu + 2*sd(Yt),
                             lower = mu - 2*sd(Yt))

randwalk_plot + geom_point(aes(y = randwalk2$mu), color = "red") + geom_point(aes(y = randwalk2$upper), color = "blue") +
                  geom_point(aes(y = randwalk2$lower), color = "blue")

```

## Done!

Congratulations!

