---
title: "Problem 3 (6 credits)"
subtitle: "HW2"
author: "Danny Moncada (monca016)"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

```{r}
suppressWarnings(suppressPackageStartupMessages({
  library(TSA)
  library(forecast)
  library(ggplot2)
  library(dplyr)
}))
```

# Boston Crime Data Analysis

## Question 1

Please pull out the crime frequency data we got from Homework 1 - Problem 3 - Question 4. You may re-plot the time series to refresh yourself about the pattern.

```{r}
setwd("~/MSBA 2020 All Files/Spring 2020/MSBA 6430 - Advanced Issues in Business Analytics/HW2")

crime=read.table("crime.txt",header=T)

N=dim(crime)[1]
crime_aggr=aggregate(rep(1,N),list(year=crime[,1],month=crime[,2],day=crime[,3]),sum)
crime_aggr=crime_aggr[order(crime_aggr[,1],crime_aggr[,2],crime_aggr[,3]),]

Y = crime_aggr[,4]

ts.plot(Y, ylab = "Frequency", main = "Boston Crime From June 2015 to September 2018")
```

### a) (1 credit)

First, let's fit an `auto.arima()` to find out a good ARIMA model for the data. Again, notice that, `auto.arima()` provides a "good" model but not necessarily the optimal. We will learn more concrete model selection techniques in Lecture 6.

**Hints**:

  - use `auto.arima()` function 

```{r}
arima_fit <- auto.arima(Y)
arima_fit
```

### b) (2 credits)

What's the model? For example
$$
(Y_t - 10) = 0.4\cdot (Y_{t-1} - 10) + e_t - 0.8 \cdot e_{t-1}
$$

**Hints**:

  - The mean value comes with every $Y_t$. In the example above, the mean value is 10.
  
  - R assumes positive sign for MA models. In the example above, R would show -0.8, rather than +0.8 for the MA(1) coefficient

Please write down the model below:
$$
Y_t = 0.99\cdot (Y_{t-1} - 270.64) + e_t - 0.71 \cdot e_{t-1} - 0.25 \cdot e_{t-2} + 0.04 \cdot e_{t-3} + 270.64
$$

### c) (1 credit)

Are any of the coefficients significant?

**Hints**:

  - A coefficient is significant if its magnitude is (roughly) at least twice as large as its standard error.


```{r}
##         ar1      ma1      ma2     ma3      mean
##      0.9888  -0.7142  -0.2542  0.0446  270.6409
##s.e.  0.0054   0.0298   0.0347  0.0292

"The first, second, and third coefficients are significant."
"The standard error*2 for ar1 is 0.0108, while the coefficient is 0.9888."
"The standard error*2 for ma1 is 0.0596, while the coefficient is 0.7142."
"The standard error*2 for ma2 is 0.0694, while the coefficient is 0.2542."
```

### d) (2 credits)

Please superimpose the fitted values on the original crime frequency time series. Does the model sufficiently explain the data?

**Hints:**

  - The fitted values can be calculated by `the original time series - arima_fit$residuals`

```{r}
fitted_vals = Y - arima_fit$residuals

par(mfrow=c(1,1))
ts.plot(Y, ylab = "Frequency", main = "Boston Crime From June 2015 to September 2018")
lines(fitted_vals,col="red")

"The auto arima model DOES sufficiently explain the data - it generally follows the process laid out by the red line."
"It doesn't capture the fringe points but not every model can be perfect."
```
