---
title: "Problem 2 (10 credits)"
subtitle: "HW2"
author: "Michael Brucek (mbrucek), Eduardo Chavez (echavezh), Kevin Grady (grady133), Danny Moncada (monca016), Tian Zhang (zhan7003)"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

```{r}
suppressWarnings(suppressPackageStartupMessages({
  library(TSA)
  library(forecast)
  library(ggplot2)
  library(dplyr)
}))
```

# Characteristic Polynomials

## Question 1

Assume $Y_t$ is the following stochastic process such as

$$
Y_t = 2.2\cdot Y_{t-1} - 1.57\cdot Y_{t-2} + 0.36\cdot Y_{t-3} + e_t
$$
where $e_i\sim N(0,1)$ i.i.d

### a) (1 credit)

First, let's determine whether the process is stationary or not by computing the roots of the characteristic polynomial.

**Hints**:

  - use `polyroot()` function 

Please pick the smallest root as $x_1$ and the larger root as $x_3$:

```{r}
x_1 <- 1.111111-0i
x_2 <- 1.250000+0i
x_3 <- 2.000000-0i
```

(please include only the numerical answer above not the computation. An acceptable format for your answer is such as `x_1 <- 5`)

### b) (1 credit)

Based on your answer above conclude whether the process is stationary or not:

```{r}
stationary <- TRUE 
# All roots are either >1 or <-1
```

### c) (2 credits)

Please generate $N=100$ sample paths of length $T=100$ for this stochastic process.

* Please save the results into a data.frame `df2c` where:
    - column `df2c$Y` has the values of the process
    - column `df2c$id` has the id of the sample path
    - column `df2c$t` has the time

```{r}
set.seed(42) # Please do not change the seed

N <- 100L
T <- 100L

ar_constant <- function(t) {
  Y = rnorm(n = t)
  for(i in 4:t) {
    Y[i] = 2.2*Y[i-1] - 1.57*Y[i-2] + 0.36*Y[i-3] + rnorm(1)
  }
  return(Y)
}

df2c <- data.frame(id = rep(1:N, each = T), t = rep(1:T, N)) %>% 
                          group_by(id) %>% 
                            mutate(Y = ar_constant(T))
```

### d) (1 credit)

Please plot the sample paths that you generated in the previous question

* Please save your plot into variable `p1d`

**Hints:**

  - use `ggplot` and take advantage of the long format of the data
  - please don't change the color (keep the lines black) but do put `alpha=0.05` into your `geom_line` to make sample paths somewhat transparent.
  - do not use `geom_points` just `geom_line` is fine
  
* As you will see from your plot:
    - the fainter the line the less likely the stochastic process would reach this spot

```{r}
p1d <- ggplot(df2c, aes(x = t, y = Y)) + geom_line(alpha = 0.5) +
          ggtitle("AR(3) Stationary Process: Y_t = 2.2*Y{t-1} - 1.57*Y{t-2} + 0.36*Y{t-3} + e_t ")
p1d
```



## Question 2 (5 credits)

Repeat a) - d) in Question 1 for the following stochastic process $Y_t$:

$$
Y_t = 2.4\cdot Y_{t-1} - 1.55\cdot Y_{t-2} + 0.3\cdot Y_{t-3} + e_t
$$
where $e_i\sim N(0,1)$ i.i.d

Compared with Question 1, we expect to see significant difference in the stationarity from the plot, although the coefficients are very close.


### a) 

First, let's determine whether the process is stationary or not by computing the roots of the characteristic polynomial.

**Hints**:

  - use `polyroot()` function 

Please pick the smallest root as $x_1$ and the larger root as $x_3$:

```{r}
x_1 <- 0.6666667+0i
x_2 <- 2.0000000-0i
x_3 <- 2.5000000+0i
```

(please include only the numerical answer above not the computation. An acceptable format for your answer is such as `x_1 <- 5`)

### b) 

Based on your answer above conclude whether the process is stationary or not:

```{r}
stationary <- FALSE 
## All roots are NOT >1 or <-1, as x_1 = 2/3
```

### c) 

Please generate $N=100$ sample paths of length $T=20$ for this stochastic process.

* Please save the results into a data.frame `df2c` where:
    - column `df2c$Y` has the values of the process
    - column `df2c$id` has the id of the sample path
    - column `df2c$t` has the time

```{r}
set.seed(42) # Please do not change the seed

N <- 100L
T <- 20L

ar_constant <- function(t) {
  Y = rnorm(n = t)
  for(i in 4:t){
    Y[i] = 2.4*Y[i-1] - 1.55*Y[i-2] + 0.3*Y[i-3] + rnorm(1)
  }
  return(Y)
}

df2c <- data.frame(id = rep(1:N, each = T), t = rep(1:T,N)) %>% 
              group_by(id) %>%
                mutate(Y = ar_constant(T))
```

### d)

Please plot the sample paths that you generated in the previous question. You should see the effect of the roots of the polynomial on the sample paths of the process.

* Please save your plot into variable `p1d`

**Hints:**

  - use `ggplot` and take advantage of the long format of the data
  - do not use `geom_points` just `geom_line` is fine
  

```{r}
p1d <- ggplot(df2c, aes(x = t, y = Y)) + geom_line(alpha = 0.5) +
  ggtitle("AR(3) Process: Y_t = 2.4*Y{t-1} - 1.55*Y{t-2} + 0.3*Y{t-3} + e_t ")
p1d
```
